---
title: "Data import & descriptives"
title-block-banner: "#870052"
title-block-banner-color: "#FFFFFF"
author: 
  name: Magnus Johansson
  affiliation: RISE Research Institutes of Sweden
  affiliation-url: https://ri.se/shic
  orcid: 0000-0003-1669-592X
date: 2023-06-12
date-format: YYYY-MM-DD
format: 
  html:
    toc: true
    toc-depth: 3
    toc-title: "Table of contents"
    embed-resources: true
    standalone: true
    page-layout: full
    mainfont: 'Arial'
    code-overflow: wrap
    code-tools: true
    code-fold: show
    code-link: true
    number-sections: true
    fig-dpi: 96
    layout-align: left
    linestretch: 1.6
    theme: materia
execute:
  echo: true
  warning: false
  message: false
  cache: false
css: style.css
editor_options: 
  chunk_output_type: console
---

## Setting up for data analysis {background-color="#99D9DD"}

Let's load packages/libraries.

```{r}
# these are mostly for data management/wrangling and visualization
library(tidyverse) # for most things
library(foreign) # for reading SPSS files
library(readxl) # read MS Excel files
library(showtext) # get fonts
library(glue) # simplifies mixing text and code in figures and tables
library(arrow) # support for efficient file formats
library(grateful) # create table+references for packages used in a project
library(styler) # only a one-time installation (it is an Rstudio plugin)
library(car) # for car::recode only
library(skimr) # data skimming
library(lubridate) # for handling dates in data
library(janitor) # for many things in data cleaning

# these are mostly for data analysis and visualization
library(psych)
library(lme4)
library(nlme)
library(broom.mixed)
library(ggplot2)
library(patchwork)
library(easystats)
library(GGally)
library(mice)
library(modelsummary)
library(ggside)
library(ggdist)
library(kableExtra)
library(formattable)
library(ggrepel)

source("ki.R") # this reads an external file and loads whatever is in it
```

### Adaptions {.smaller}

Some functions exist in multiple packages, which can be a source of headaches and confusion.

Below we define preferred functions that are frequently used. If desired, we can still use specific functions by using their package prefix, for instance `dplyr::recode()`.

```{r}
### 
select <- dplyr::select
count <- dplyr::count
recode <- car::recode
rename <- dplyr::rename
```

## Correlation & visualization

rework these for new dataset, first without "styling"

### Correlation matrix

without and with grouping

```{r}
#| eval: false
df %>% 
  select(starts_with("WHO"),GDS,WHOQoL_BREF,NMM,Gender) %>% 
  ggpairs(aes(color = Gender),alpha = 0.8) +
  scale_color_manual(values = RISEpalette1,
                     aesthetics = c("color", "fill"))
```

### easystats `plot(cor_test())`

```{r}
#| eval: false
plot(cor_test(df, "NMM", "WHOQoL_BREF")) +
  theme_minimal(base_size = 15, 
                base_family = "Lato") +
  theme_rise(fontfamily = "Lato", 
             axissize = 15) +
  geom_point(data = df, 
             aes(NMM, WHOQoL_BREF), 
             size = 2) +
    ylab("WHOQoL-BREF") +
    xlab("NMM")
```
