---
title: "Data import & descriptives"
title-block-banner: "#870052"
title-block-banner-color: "#FFFFFF"
author: 
  name: Magnus Johansson
  affiliation: RISE Research Institutes of Sweden
  affiliation-url: https://ri.se/shic
  orcid: 0000-0003-1669-592X
date: 2023-06-12
date-format: YYYY-MM-DD
format: 
  html:
    
    toc: true
    embed-resources: true
    standalone: true
    mainfont: 'Arial'
    code-overflow: wrap
    code-fold: show
    code-link: true
    number-sections: true
    fig-dpi: 96
    layout-align: left
    linestretch: 1.6
    theme: materia
  pdf:
    papersize: a4
    documentclass: article #article, report or book
    #classoption: [twocolumn, portrait]
  docx:
    
    toc: false
    number-sections: false
    highlight-style: github
always_allow_html: true
execute:
  echo: true
  warning: false
  message: false
  cache: false
css: style.css
editor_options: 
  chunk_output_type: console
---

## Overview of this file

- Visualizing distributions
- Running models
- Checking assumptions
- Effects(?)

### Setting up

Let's load packages/libraries.

```{r}
#| code-fold: show
# these are mostly for data management/wrangling and visualization
library(tidyverse) # for most things
library(foreign) # for reading SPSS files
library(readxl) # read MS Excel files
library(showtext) # get fonts
library(glue) # simplifies mixing text and code in figures and tables
library(arrow) # support for efficient file formats
library(grateful) # create table+references for packages used in a project
library(styler) # only a one-time installation (it is an Rstudio plugin)
library(car) # for car::recode only
library(skimr) # data skimming
library(lubridate) # for handling dates in data
library(janitor) # for many things in data cleaning

# these are mostly for data analysis and visualization
library(gtsummary)
library(scales)
library(visdat)
library(psych)
library(lme4)
library(nlme)
library(broom.mixed)
library(patchwork)
library(easystats)
library(mice)
library(modelsummary)
library(ggdist)
library(kableExtra)
library(formattable)
library(ggrepel)
library(ggrain)
```

Define a ggplot theme `theme_ki()`, a standard table function, `kbl_ki()`, and a color palette based on KI's design guide, `ki_color_palette`.

```{r}
#| code-fold: show
source("ki.R") # this reads an external file and loads whatever is in it
```


## Visualizations

### Distributions

#### Histogram

```{r}
df.long %>% 
  filter(Group == "Control") %>% 
  ggplot(aes(x = value, fill = measure)) +
  geom_histogram(binwidth = 4) +
  facet_grid(measure~time) +
  theme_ki() +
  scale_fill_manual(values = ki_color_palette)

```

#### Density

```{r}
df.long %>% 
  filter(Group == "Control") %>% 
  ggplot(aes(x = value, fill = measure)) +
  geom_density() +
  facet_grid(measure~time) +
  theme_ki() +
  scale_fill_manual(values = ki_color_palette)
```

What do these distributions look like to you? Normal?

### Individuals over time

```{r}
df.long %>%
  filter(measure == "DEP") %>% 
  ggplot(aes(x = time, y = value, group = time, fill = Group, color = Group)) +
  geom_rain(
    #alpha = .6,
    boxplot.args = list(color = "black", outlier.shape = NA),
    id.long.var = "id"
  ) +
  scale_fill_brewer(palette = "Dark2",
                    aesthetics = c("color","fill")) +
  theme_ki() +
  facet_wrap(~Group,
             nrow = 2)
```


## Linear model

summary 
glance 
tidy 
tbl_
tab_model?

```{r}
pred2 <- estimate_expectation(m2, data = "grid")
plot(pred2) +
  scale_x_continuous(breaks = c(0,1)) +
  theme_minimal() +
  theme_rise()
plot(parameters(m2))
```

### Check assumptions



link to `marginaleffects`!

meme

https://twitter.com/stephenjwild/status/1666056019993034755
